Constraint Grammar as a SAT problem

### 2015-06-01 paper/nodalida2015.tex
stuff

The sequential application of traditional CG rules is good for
performance and transparency. As soon as a rule takes action, the analyses
are removed from the input, and next rules get the modified sentence
as their input. 
% At the execution of rule, there is no way to go back to earlier rules and undo them. 
Thus there is no way to know which part comes directly from the
raw input and which part from applying previous rules.

A conflict in an ordered scheme can be defined as 
a set of two or more rules, such that applying the first makes the
next rules impossible to apply, regardless of input.
We can reuse the example from Section~\ref{sec:unord}:

@@

The first rule selects the verb always, removing all other readings,
and the second rule removes verb reading, given certain condition.
If the rules are introduced in a different order, there is no
conflict: the REMOVE rule would not remove verb readings from all
possible verb analyses, so there is a possibility for the SELECT rule
to fire.

Ordered SAT-CG can be used to detect these conflicts without any
modifications, as a side effect of its design. 
After applying each rule, it stores the clauses
produced by it and commits to them
% ---even if the new rules cannot
% change the earlier decisions, the SAT solver still retains the
% dependencies.
% Given that each application of a rule 
% is 
% gets a variable,
In case of a conflict, the program detects the particular rule that 
violates the previous clauses, with the sentence
where it is applied. Thus we get feedback which rule fails, and on which
particular word(s).

This conflict solving can be done per section, or for the whole grammar. 
In case of section-based application, we run each section on the
input, and give the output as a fresh input to the new section, discarding
all clauses. Alternatively, we can let the rules interact with each
other regardless of the section boundaries.

The unordered scheme with maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
The whole definition of conflicts depends on ordering, and it would
not differentiate between the case where \texttt{SELECT v} comes
before \texttt{REMOVE v IF (-1 det)} or after.
On the other hand, an unordered formalism such as Finite-State
Intersection Grammar \cite{koskenniemi90} would benefit from the
maximisation-based technique in conflict handling.

>>>>

The sequential application of traditional CG rules is good for
performance and transparency. When a rule takes action, the analyses
are removed from the input, and the next rules get the modified sentence
as their input. 
% At the execution of rule, there is no way to go back to earlier rules and undo them. 
As a downside, there is no way to know which part comes directly from the
raw input and which part from applying previous rules.

A conflict in an ordered scheme can be defined as 
a set of two or more rules, such that applying the first makes the
next rules impossible to apply, regardless of the input.
We can reuse the example from Section~\ref{sec:unord}:

@@

The first rule selects the verb reading always, removing all other readings,
and the second rule removes the verb reading conditionally.
If the rules are introduced in a different order, there is no
conflict: the REMOVE rule would not remove verb readings from all
possible verb analyses, so there is a possibility for the SELECT rule
to fire.

Ordered SAT-CG can be used to detect these conflicts without any
modifications, as a side effect of its design. 
After applying each rule, it stores the clauses
produced by the rule and commits to them.
In case of a conflict, the program detects the particular rule that 
violates the previous clauses, with the sentence
where it is applied. Thus we get feedback which rule fails, and on which
particular word(s).

% This conflict solving can be done per section, or for the whole grammar. 
% In case of section-based application, we run each section on the
% input, and give the output as a fresh input to the new section, discarding
% all clauses. Alternatively, we can let the rules interact with each
% other regardless of the section boundaries.

Unordered SAT-CG with maximisation-based conflict solving is not
suitable for this task: the whole definition of conflict depends on
ordering, and the unordered scheme deliberately loses this information.
On a more speculative note, an unordered formalism such as Finite-State
Intersection Grammar \cite{koskenniemi90} might benefit from the
maximisation-based technique in conflict handling.
